---
draft: false
title: Methodolody and Results
weight: 30
---
```{r, echo = FALSE, message = FALSE,  warning=FALSE, results='hide'}
source(here::here("scripts/setup.R"))
source(here::here("scripts/Data_cleaning_script.R"))
source(here::here("scripts/EDA_script.R"))
source(here::here("scripts/Analysis_script.R"))
```

We report here the results which are worth mentioning in our opinion. As said above, we select the fixed effect method. 
The model is:

$$
\begin{align*}
Y_{i,t} &=\alpha+\beta_1log(GDP)+\beta_2mh\_exp\_pc+\beta_3perc\_bscholder\_25\_44+\\ & \,\,+\beta_4 White+\beta_5 BlackAfricanAmerican+\beta_6Asian + \\
& \,\, + \beta_7Age\_0\_17 + \beta_8Age\_18\_24+ \beta_9Age\_25\_44+ \\ & \,\,+\beta_{10}Age\_45\_64+ \beta_{11}Age\_65\_84+\beta_{12}log(population)
\end{align*}
$$
$Y_{i,t}$ refers to the dependent variable for state $i$ at time $t$.
The estimation is done considering $Y_{i,t}-\bar{Y_i}$, where $\bar{Y_i}$ is the mean dependent variable for the state $i$. Indeed, $\alpha$ will not appear among the estimates, as it is constant overtime and cancels out.

For total criminality regression's results are:

```{r}
tab_model(fe_tot, title="Fixed Effect, dependent variable: Total Criminality per 1000 inhabitants")
```

We can notice that the $R^2$, which is a statistical measure representing the proportion of the variance for a dependent variable that's explained by independent variables in a regression model, is lower here with respect to the standard OLS.
With respect to the standard OLS estimations, magnitudes changes but not of sign. The only exception is mental health expenditure which, here, appears having a positive effect on criminality. Although, mh_exp_pc and education's proxy are not statistically significant anymore. 
Additionally, among races, only the percentage of asian in the population seems statistically significant and still negative influencing criminality.
As in the OLS estimates, $log(population)$ decreases criminality: as population increases by 1%, criminality decreases by 16 crimes per 1000 inhabitants circa.

Among the various regressions we run, only the ones with rape and homicide as dependent variables have different results from the one just presented above. 

For Rape:

```{r}
tab_model(fe_rape, digits=4, title="Fixed Effect, dependent variable: Rapes per 1000 inhabitants")
```

From the FE regression with Rape per 1000 inhabitants as dependent variable we learn that:

* Rape appears, in magnitude, less impacted by GDP and population than whole criminality, but still respectively positively and negatively,
* $R^2$ is very low, so probably the regression does not explain the variance in Rape in a state in a certain year,
* Only GDP and $log(Population)$ are statistically significant, the other variables' estimated coefficients are not.

For Homicides:

```{r}
tab_model(fe_hom, digits=3, title="Fixed Effect, dependent variable: Homicides per 1000 inhabitants")
```

From the FE regression with Homicides per 1000 inhabitants as dependent variable we learn that:

* Homicides appear, in magnitude, less impacted by GDP and population than whole criminality, but still respectively positively and negatively,
* $R^2=0.683$, so the variables in the regression explain 68% of the variance in homicides per 1000 people in a state in a certain year,
* Black African American and Asian percentages in the population have a, respectively, statistically significant positive and negative effect on criminality per 1000 terms.
* Having a high percentage of very young population appears to increase homicides at 5% significance, but this is difficult to explain through social mechanism in a community for us.