---
title: "Crime data"
draft: false
weight: 15
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
---

```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/setup.R"))
source(here::here("scripts/Data_cleaning_script.R"))
```

### Sources and Description

We used the dataset on estimated crimes (from 1979 to 2019) available in the [FBI website](https://crime-data-explorer.fr.cloud.gov/downloads-and-docs). We have repeated observations for each state in the United States of America from 1979 to 2019. 

```{r, echo = FALSE, message = FALSE, warning=FALSE}
 DT::datatable(estimated_crimes_1979_2019, options = list( scrollX='400px', scrollY='300px'), caption = 'Raw data - Estimated Crimes 1979-2019 in US' )
```

The data-set contains 2116 observations for 15 variables, which are

<ul>
  <li> <b>year</b> the year of the observation </li>
  <li> <b>state_abbr</b>, <b>state_name</b> the abbrevation and the name of the
  State. Notice that the first line of each   year is blank. These observations
  refer to the total, i.e. United States.
  </li>
  <li><b>population</b> the number of population in a given year and State.
  </li>
  <li> <b>violent_crime</b>, <b>homicide </b>, <b>rape_legacy</b>,
  <b>rape_revised</b>, <b>robbery</b>, <b>aggravated_assault</b>,
  <b>property_crime</b>, <b>burglary</b>, <b>larceny</b>, <b>motor_vehicle_theft</b>, <b>violent_crime</b>, <b>caveats</b>
  each refers to the respective number of registered crimes by the FBI or of caveats. From the source we learn that homicides resulting from the events of September 11, 2001, are not included. This is fine for us, because it will be an outlier not significant for our analysis. 
  </li>
</ul>
<br>

```{r}
kable(as.data.frame(colSums(is.na(estimated_crimes_1979_2019))), 'simple', col.names = "Missing Values", caption='Missing values for each feature')
```

From above we can see how many NAs we have for each feature. Looking at this we already decide to not take into account <b>rape_revised</b> and <b>caveats</b>, while we already now that <b>state_name</b> and <b>state_abbr</b> missing values refers to United States, so we will fill them appropriately.

### Wrangling/cleaning

To clean this dataset we have to transform year values into numeric. Moreover, we change the name of column <b>state_name</b> to <b>State</b> and selected only some crime which we think could be more relevant for our study and could be more impacted by mental health expenditure.
We also replaced NAs in <b>State</b> and <b>state_abbr</b> with "United States" and "US".

The cleaned dataset is called "estimated_crimes" and is reported below:
<br>

```{r, echo = FALSE, message = FALSE, warning=FALSE}
  DT::datatable(estimated_crimes, options = list(scrollX='400px', scrollY='300px'), caption = 'Cleaned data - Estimated Crimes 2004-2013 in US')
```
