---
title: "US Demographics"
draft: false
weight: 35
---
```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/setup.R"))
source(here::here("scripts/Data_cleaning_script.R"))
```

### Sources and Description

We used 3 data-sets from the [United States Census Bureau's website](https://www.census.gov/){target="_blank"}.  

The first one is about race composition from 2000 to 2010
```{r, fig.cap='Raw data - Demographics, Race, 2000-2010', echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(st_est00int_race, options = list( scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```

This one contains 364 observations for 18 variables, which are

<ul>
  <li> <b>REGION</b>,<b>DIVISION</b>, <b>STATE</b>, <b>NAME</b> which identifies the region, division, state code and name of the state. </li>
  <li> <b>RACE</b> is the race, it goes from 0 to 6, with 0 being the total and 1-6 identifying ethinicities as White, Black/African-American and so on. </li>
  <li> <b>POPBASE2000</b> and <b>POPESTIMATExxxx</b> for each year, with "xxxx" being the referred year, which are the estimated population in a State in a given year, for the respective race cathegory.</li>
</ul>

The second one is about age and sex composition from 2000 to 2010

```{r, fig.cap='Raw data - Demographics, Age & Sex, 2000-2010', echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(st_est00int_agesex, options = list(scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```
This one contains 13572 observations for 19 variables, which are

<ul>
  <li> <b>REGION</b>,<b>DIVISION</b>, <b>STATE</b>, <b>NAME</b> which identifies the region, division, state code and name of the state. </li>
  <li> <b>SEX</b> is the sex, it can be either 0 (total), 1 (male) or 2 (female). </li>
  <li> <b>AGE</b> is the age, it goes from 0 to 85 years old, then there's also 999 which is the total population. </li>
  <li> <b>POPBASE2000</b> and <b>POPESTIMATExxxx</b> for each year, which are the estimated population in a State in a given year, for the respective sex and age cathegory.</li>
</ul>

The third one is about race, age and sex composition from 2010 to 2019

```{r, fig.cap='Raw data - Demographics, Race, Age & Sex, 2010-2019', echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(st_est00int_agesex, options = list(scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```
This one contains 236844 observations for 21 variables, which are

<ul>
  <li><b>SUMLEV</b> is the identification of the summary levels used by the census, it is also called "area type".</li>
  <li> <b>REGION</b>,<b>DIVISION</b>, <b>STATE</b>, <b>NAME</b> which identifies the region, division, state code and name of the state. </li>
  <li> <b>RACE</b> is the race, it goes from 1 to 6, 1-6 identifying ethinicities as White, Black/African-American and so on </li>
  <li> <b>ORIGIN</b> is the origin, it can be 0 (total), 1 (Not Hispanic) or 2(Hispanic), although this information is absent between 2000 and 2009 so we will omit it. </li>
  <li> <b>SEX</b> is the sex, it can be either 0 (total), 1 (male) or 2 (female). </li>
  <li> <b>AGE</b> is the age, it goes from 0 to 84 years old, then we have 85 which comprises 85+ years old </li>
  <li> <b>POPBASE2010</b> and <b>POPESTIMATExxxx</b> for each year, which are the estimated population in a State in a given year, for the respective race, sex and age cathegory. </li>
</ul>


### Wrangling/cleaning

The cleaning is done for each data-set separately. Later we proceed to join them.
 In all data-set we trasform REGION into a factor and we rename the levels such that total (US) replaces 0, North-East (NE) replaces 1, Mid-West (MW) replaces 2, South (S) replaces 3, and West (W) replaces 4. 

Also RACE and SEX will become factors  with respective levels labels: 
(White = 1, BlackAfricanAmerican = 2, AmericanIndianAlaska = 3, Asian = 4, HawaiianPacificIslanders = 5, Racegreaterthan1 = 6 and, for the dataset in years 2000-2010, Total = 0) and (Total=0, Male=1, Female=2).

Names of variables are also changed slightly to have them in line with other data-sets and through pivot_longer and pivot_wider we adjust the structure of the table in a standardized way. 

Moreover, for AGE we create some sub-groups instead of having the complete range 0-85 years old. 
The age groups we create are 0-17, 18-24, 25-44, 45-64, 65-84 and 85+. We still don't know whether age composition has an impact on criminality, but we consider important to have the subgroups 18-24 and 25-44, since in education, as we will see, these two age groups are considered. 

Also race groups are different in the cleaned dataset: White, BlackAfricanAmerican, Asian and Other_race. The latter comprises all the other cathegories.
We also filter for years of interest (2004-2013).

In 2011-2013 we miss the observation for United States, which instead is present between 2004-2010. Therefore, we creat a dataset for it by taking the sum across states, since US's values would be the total and we put everything together for years 2011-2013 to obtain the whole data-set in this time framework.

We end up having two data-sets on demographics, one for the years 2004-2010 and the other from 2011 to 2013. 
Finally, we join these obtaining the final "demographics" dataset:

```{r,  fig.cap='Cleaned data - Demographics in US', echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(demographics, list(scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```

No missing value is present. Although, notice that in the cleaning process we have to fill some Region's values which otherwise would be missing. But, knowing the data-set and the State, it is straightforward.