---
title: "Gross Domestic Product"
draft: false
weight: 25
---
```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("scripts/setup.R"))
source(here::here("scripts/Data_cleaning_script.R"))
```

### Source and Descriptions

The data-set on GDP can be found in the [Bureau of Economic Analysis, of U.S. Department of Commerce, website](https://apps.bea.gov/iTable/iTable.cfm?7001=1200&7002=1&7003=200&7004=sic&7005=1&7006=xx&7007=-1&7035=-1&7036=-1&7090=70&7093=levels&isuri=1&reqid=70&step=10).

```{r, fig.cap="Raw Data - Dataset which gives us info on GDP, our variable of interest", echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(GDP, list(scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```

This has 483 observations for 27 columns, which are:
<ul>
  <li> <b>GeoFips</b> and <b>GeoName</b> which identify the state, or the whole US through a code and the name, respectively. </li>
  <li><b>LineCode</b> and <b>Description</b> which identify the kind of variable we are looking at, i.e. Current-dollar GDP (millions of current dollars), Real GDP (millions of chained 2012 dollars), etc... They are of 8 different kind, but we will focus on Current-dollar GDP as you will see.  </li>
  <li> <b>1997 ... 2019</b> one column for each year observed. </li>
</ul>

### Wrangling/Cleaning

In order to clean the dataset we filter for one values of <b>Description</b> only, since it's the one of interest for us: Current-dollar GDP (millions of current dollars). We delete the column which are not relevant, remaining with renamed <b>GeoName</b>, which is now <b>State</b>, and <b>1997 ... 2019</b>. We use "pivot_longer" to create two variables <b>year</b> and <b>Current_dollar_GDP_millions</b> increasing the length of the data-set. 
Of course, we also filter for years in the time-frame 2004-2013.

The resulting data-set is "GDP_Cleaned":

```{r, fig.cap="Cleaned Data - Current-dollar GDP (millions of current dollars)", echo = FALSE, message = FALSE, warning=FALSE}
DT::datatable(GDP_Cleaned, list(scrollX='400px',
  pageLength = 5,
  lengthMenu = c(5, 10, 15, 20)))
```

No missing values are present in "GDP_Cleaned".